
---
title: "Choropleth Map for Minnesota Counties Population in 2010"
author: "Long Truong"
output:
  html_document
    
---

## Load packages

Download the following libraries (ggplot2, sf, dplyr, tidylog, and ggiprah) and load them.
```{r, message=FALSE}
library(ggplot2)
library(sf)
library(dplyr)
library(tidylog)
library(ggiraph)
```

## Data Wrangling

### Step 1
Download the *Shapefile* data from [Minnesota Geospatial Commons](https://gisdata.mn.gov/dataset/bdry-counties-in-minnesota). The page was found via a Google search using the term `Minnesota Shapefile`. 
Use the `read_sf` function from the `sf` library to capture the data to a table. We will call it `mn`.
```{r, message=FALSE}
mn <- read_sf("./mn_shapefile_2/", "mn_county_boundaries_500")
head(mn)
```

### Step 2
* Next, get the Minnesota population data in year 2010 to get more features like population, households, and persons per household. \
* From the [Popfinder for Minnesota Counties](https://mn.gov/admin/demography/data-by-topic/population-data/our-estimates/pop-finder1.jsp), press *Download all the data in this PopFinder tool. (Excel file)* \
* Move the Excel file to the working folder (meaning the same folder as your R file) and name it to `mn_county_bigdata.cvs`.

### Step 3

Clean the data by just considering the county geography type and the year 2010 from the data.

```{r, message=FALSE}
mn_clean <- read.csv("mn_county_bigdata.csv")
mn_clean <- mn_clean %>% 
  filter(Year==2010 & Geography.Type=="County")
head(mn_clean)
```

Perform a left join to to our main data `mn` so it will have the additional features.
```{r, message=FALSE}
# Add Households and Persons Per Household variables to the main data table
mn <- mn %>% left_join(mn_clean, by = c("CTY_NAME" = "Geography.Name"))
```

### Step 4

Now, let's create categories for the population from very small to very large. 

```{r, message=FALSE}
# Define the breakpoints for the categories for population
breakpoints <- c(0, 5000, 10000, 50000, 100000, Inf)


# Convert to integer
mn$Pop_int <- gsub(",", "", mn$Population) %>%
  as.integer()

# Create categories
mn$pop_category <- cut(mn$Pop_int, breaks = breakpoints, labels = c("very small", "small", "medium", "large", "very large"), include.lowest = TRUE)

```

This code chunk below combines the wanted features of a county into a character string. This helps as when you hover over a county on the map, it displays the correct information of name, population, households, and persons per household.

```{r, message=FALSE}
mn <- mn %>%
  mutate(info = paste(
    "Name: ", CTY_NAME,
    "\nPopulation Category: ", pop_category,
    "\nPopulation: ", Population,
    "\nHouseholds: ", Households,
    "\nPersons Per Household: ", Persons.Per.Household..PPH.
  ))
```


## Step 5

Let's plot a static choropleth map showing Minnesota counties by population in 2010 first. We use ggplot with passed in parameters below. 

```{r, fig.alt="Choropleth map showing Minnesota counties by population in 2010" }
plot <- ggplot(mn, aes(fill = pop_category)) +
  geom_sf() + # call geom shapefile
  geom_sf_interactive(aes(geometry = geometry, tooltip = info)) + 
  # this helps with the future interactive feature
  scale_fill_manual(
    values = c(
      "very small" = "#eff3ff",
      "small" = "#bdd7e7",
      "medium" = "#6baed6",
      "large" = "#3182bd",
      "very large" = "#08519c") ) # color scheme to match the category +
  labs(title = "Minnesota counties by population category in 2010") +
  theme(legend.position = "right")

plot
```

## Final Product
Finally, we use this code below to produce the interactive map

```{r,fig.alt="Interactive choropleth map showing Minnesota counties by population in 2010"}
final <- ggiraph(code = print(plot))  
final
```



## Resource used

[GGiraph Interactive Map Documents](http://davidgohel.github.io/ggiraph/) 
\
[Minnesota Geospatial Commons](https://gisdata.mn.gov/dataset/bdry-counties-in-minnesota)
\
[Additional Minnesota Population data](https://mn.gov/admin/demography/data-by-topic/population-data/our-estimates/pop-finder1.jsp) \

